Hippocampal replay of experience at real-world speeds

### 2020-10-19 Article.tex

In the stationary movement dynamic, the latent position does not change between time steps. The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position (subject to the size of the position bin). In the continuous movement dynamic, the latent position is most likely to be  "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix. This means that, for a given latent position, the probability of moving to another position is modeled by a Gaussian centered at that position and "spatially close" is defined by the variance of the Gaussian. In our case, since replay has been shown to move at speeds much faster than the animal's speed \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}, we set the variance of the Gaussian to 6.0 cm. This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or equivalently, this means that latent speeds from 0 to ~25 m/s are most likely).  Finally, in the fragmented movement dynamic, the latent position can move to any available position instantaneously. We model this using a uniform state transition matrix, which makes transitions to each position equally likely.

@@

The second most prevalent classification was exclusively stationary events (5149 of 23382 or 22\% classified SWRs). Unlike the stationary-continuous mixtures, most of these events represented a location close to the animal's actual position (Figure \ref{5}C). There were, however, a number of stationary events that represented positions far from the animal. We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in 1551 of 23382 (~7\%) of classified SWRs. About half of these (731 classified SWRs) were stationary throughout the duration of the SWR. These non-local stationary events most commonly represented reward wells or choice points (Figure \ref{5}E), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track. This suggests that these representations can be flexibly deployed to represent discrete locations.

@@

Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of <1 m/s. The next most common category were stationary events \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events ---continuous, extended trajectories through space---made up only about 19\% of classified events. Importantly, events with slower dynamics tended to persist for approximately the same length of time as events (\texttildelow50-100 ms) with faster continuous dynamics, consistent with a brief activation of a representation of a location or a short trajectory through space. Events with slower dynamics could also represent either the animal's current location or locations far from the animal, and similarly, previous work has shown that events with continuous dynamics most often begin at the animal's current location (\cite{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009})

>>>>

In the stationary movement dynamic, the latent position does not change between time steps. The state transition matrix can thus be defined as an identity matrix, which predicts the next position will be the same as the last position (subject to the size of the position bin). In the continuous movement dynamic, the latent position is most likely to be  "spatially close" to the position in the previous time step, so we use a Gaussian random walk state transition matrix. This means that, for a given latent position, the probability of moving to another position is modeled by a Gaussian centered at that position and "spatially close" is defined by the variance of the Gaussian. In our case, since replay has been shown to move at speeds much faster than the animal's speed \cite{DavidsonHippocampalReplayExtended2009, PfeifferAutoassociativedynamicsgeneration2015}, we set the variance of the Gaussian to 6.0 cm. This ensures that with a 2 ms time step, the latent position is 95\% likely to be within 4.90 cm of the previous latent position (or equivalently, this means that latent speeds from 0 to \texttildelow25 m/s are most likely).  Finally, in the fragmented movement dynamic, the latent position can move to any available position instantaneously. We model this using a uniform state transition matrix, which makes transitions to each position equally likely.

@@

The second most prevalent classification was exclusively stationary events (5149 of 23382 or 22\% classified SWRs). Unlike the stationary-continuous mixtures, most of these events represented a location close to the animal's actual position (Figure \ref{5}C). There were, however, a number of stationary events that represented positions far from the animal. We set a threshold of 30 cm distance from the animal to identify non-local stationary events and found such content in 1551 of 23382 (\texttildelow7\%) of classified SWRs. About half of these (731 classified SWRs) were stationary throughout the duration of the SWR. These non-local stationary events most commonly represented reward wells or choice points (Figure \ref{5}E), consistent with \cite{JaiDistincthippocampalcorticalmemory2017}, but a small portion of them occurred at other locations of the track. This suggests that these representations can be flexibly deployed to represent discrete locations.

@@

Importantly, the spatially coherent events most often expressed a dynamic that was consistent with slower speeds of <1 m/s. The next most common category were stationary events \cite{JaiDistincthippocampalcorticalmemory2017, FarooqEmergencepreconfiguredplastic2019} that activated an extended representation of a single location, while the "classical" replay events ---continuous, extended trajectories through space---made up only about 19\% of classified events. Importantly, events with slower dynamics tended to persist for approximately the same length of time as events (\texttildelow50-100 ms) with faster continuous dynamics, consistent with a brief activation of a representation of a location or a short trajectory through space. Events with slower dynamics could also represent either the animal's current location or locations far from the animal, and similarly, previous work has shown that events with continuous dynamics most often begin at the animal's current location \cite{DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.
